Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-05-04 second-draft/pnas-word-order-si.tex
si

\begin{table*}[ht]
	\begin{center}
\begin{tabular}{|ll|l|l|}
	\hline
	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
	verb & object &    \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP    &    case   & 3, 4   \\ \hline
copula  verb  &    predicate    &    cop   & --    \\\hline
`want'    &    VP    &    xcomp   & 15   \\\hline
tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & 16, 13  \\
	negative auxiliary    &    VP    &    & -- \\ \hline
complementizer    &    S    &   \multirow{2}{*}{mark}  & --    \\
	adverbial subordinator & S &  & -- \\ \hline
noun    &    genitive    &    nmod & 2, 23   \\ \hline
noun    &    relative clause    &    acl   &  24    \\ \hline
	adjective & standard of comparison & \multirow{2}{*}{obl} & --\\
verb    &    PP    &    & 22   \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{3}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations according to Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies. We also provide the corresponding numbers in Greenberg's original study.}\label{table:greenberg-dryer}
\end{table*}

@@

\section{Posterior}

\begin{figure}
	\begin{center}
	\includegraphics[width=0.48\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
	\end{center}
	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
\end{figure}

@@

In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations and relative optimality of language.

The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (TODO check).

@@

>>>>

\begin{table*}[ht]
	\begin{center}
\begin{tabular}{|ll|l|l|}
	\hline
	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
	verb & object &    \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP    &    case   & 3, 4   \\ \hline
copula  verb  &    predicate    &    cop   & --    \\\hline
`want'    &    VP    &    xcomp   & 15   \\\hline
tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & 16, 13  \\
	negative auxiliary    &    VP    &    & -- \\ \hline
complementizer    &    S    &   \multirow{2}{*}{mark}  & --    \\
	adverbial subordinator & S &  & -- \\ \hline
noun    &    genitive    &    nmod & 2, 23   \\ \hline
noun    &    relative clause    &    acl   &  24    \\ \hline
	adjective & std. of comp. & \multirow{2}{*}{obl} & --\\
verb    &    PP    &    & 22   \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{3}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations according to Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies.
	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study.
	}\label{table:greenberg-dryer}
\end{table*}

@@

\section{Comparing Efficiency to its Components}

\begin{figure}[ht]
	\begin{center}
	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
	\end{center}
	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
\end{figure}

@@

In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.

The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations.
We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (TODO check).

@@

\begin{figure}[ht]
    \centering
